

```{r warning=FALSE, include=FALSE}
pacman::p_load(tidyverse)
pacman::p_load(knitr,dplyr,AICcmodavg)
pacman::p_load(inspectdf,tidyr,stringr, stringi,DT,mice)
pacman::p_load(caret,modelr)
pacman::p_load(mlbench,mplot)
pacman::p_load(tidymodels,glmx)
pacman::p_load(skimr,vip,yardstick,ranger,kknn,funModeling,Hmisc)
pacman::p_load(ggplot2,ggpubr,GGally,corrplot, ggthemes)
knitr::opts_chunk$set(message = FALSE)
```


```{r warning=FALSE}
# Load data
data <- read.csv("finalData.csv", header = TRUE, sep = ",")
nrow(data)
names(data)
# head(data)
```

```{r warning=FALSE}
data <- data[, !(names(data) %in% c("TotalClaimAmount", "OPTotalAmount", "PhysRoleCount"))]
```

```{r warning=FALSE}
# print all the columns
names(data)
```

```{r warning=FALSE}
# Check the distribution and basic information of the dataset
print("Summary of the dataset:")
summary(data)

print("Structure of the dataset:")
str(data)
```

```{r}
# Get the frequency distribution of PotentialFraud
print("Frequency distribution of PotentialFraud:")
freq <- table(data$PotentialFraud)
print(freq)
```

```{r}
# Get the percentage distribution of PotentialFraud
print("Percentage distribution of PotentialFraud:")
percentage <- prop.table(freq) * 100
print(percentage)
```

```{r}
# Convert PotentialFraud to a factor for visualization
data$PotentialFraud <- as.factor(data$PotentialFraud)

# Create a bar plot for PotentialFraud distribution
ggplot(data, aes(x = PotentialFraud)) +
  geom_bar() +
  ggtitle("Distribution of Potential Fraud") +
  xlab("Potential Fraud") +
  ylab("Count") +
  scale_y_continuous(labels = scales::comma)
```




### 2.2.2 Gender and Race
```{r}
# Convert gender and race to factors 
data$Gender <- as.factor(data$Gender)
data$Race <- as.factor(data$Race)
```

```{r}
# Check unique values
unique_values_gr <- sapply(data[,c('Gender', 'Race')], unique)
unique_values_gr

# Plot Gender and Race
gender_plot <- ggplot(data, aes(x = Gender)) +
  geom_bar() +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle('Gender Distribution')

race_plot <- ggplot(data, aes(x = Race)) +
  geom_bar() +
  theme_minimal() +
  scale_y_continuous(labels = scales::comma) +
  ggtitle('Race Distribution')

grid.arrange(gender_plot, race_plot, ncol = 1)
```

# 2.2.3 Health Conditions
```{r}
# Check unique values for health conditions
unique_values_hc <- sapply(data[,c('ChronicCond_Alzheimer', 
                                        'ChronicCond_Heartfailure', 
                                        'ChronicCond_KidneyDisease', 
                                        'ChronicCond_Cancer', 
                                        'ChronicCond_ObstrPulmonary', 
                                        'ChronicCond_Depression', 
                                        'ChronicCond_Diabetes', 
                                        'ChronicCond_IschemicHeart', 
                                        'ChronicCond_Osteoporasis', 
                                        'ChronicCond_rheumatoidarthritis', 
                                        'ChronicCond_stroke')], unique)
print(unique_values_hc)
```


# Calculate the correlation matrix
cor_matrix <- cor(data[,c('ChronicCond_Alzheimer', 
                               'ChronicCond_Heartfailure', 
                               'ChronicCond_KidneyDisease', 
                               'ChronicCond_Cancer', 
                               'ChronicCond_ObstrPulmonary', 
                               'ChronicCond_Depression', 
                               'ChronicCond_Diabetes', 
                               'ChronicCond_IschemicHeart', 
                               'ChronicCond_Osteoporasis', 
                               'ChronicCond_rheumatoidarthritis', 
                               'ChronicCond_stroke')], method="spearman")

# Create the correlation plot with coefficients displayed
corrplot(cor_matrix, method="shade", type="upper", order="hclust", 
         addCoef.col="black",  # Color of the coefficients
         number.cex=0.7,      # Font size of coefficients
         tl.cex = 0.7,        # Font size of variable names
         title="Correlation Plot of Health Conditions")



```{r}
# Convert the wide dataset to a long format for easier plotting
data_long <- data %>%
  gather(key = "ChronicCondition", value = "Value", 
         c('ChronicCond_Alzheimer', 'ChronicCond_Heartfailure', 'ChronicCond_KidneyDisease', 
           'ChronicCond_Cancer', 'ChronicCond_ObstrPulmonary', 'ChronicCond_Depression', 
           'ChronicCond_Diabetes', 'ChronicCond_IschemicHeart', 'ChronicCond_Osteoporasis', 
           'ChronicCond_rheumatoidarthritis', 'ChronicCond_stroke'))

# Ensure that the Value column is a factor
data_long$Value <- as.factor(data_long$Value)

# Plot the data
data_long %>%
  ggplot(aes(x = Value, fill = Value)) +
  geom_bar(position = 'identity') +
  coord_flip() + 
  facet_wrap(~ChronicCondition, scales = "free", ncol = 3) +
  theme_minimal() +
  theme_economist() +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank()) +
  labs(fill = "Condition Status", y = "Frequency", x = "") +
  scale_fill_manual(values = c("1" = "#312876", "0" = "#7c9c86"), 
                    labels = c("1" = "Chronic", "0" = "No Chronic"))
```


##2.2.4 Financial Details

```{r}
# Check distribution of financial columns
summary(data[,c('TotalClaimAmount', 'OPTotalAmount', 'Log_TotalClaimAmount', 'Log_IPTotalAmount', 'Log_OPTotalAmount')])

# Convert wide format to long for plotting
financial_long <- gather(data, key="FinancialType", value="Value", TotalClaimAmount, OPTotalAmount, Log_TotalClaimAmount, Log_IPTotalAmount, Log_OPTotalAmount)

# Create a histogram
ggplot(financial_long, aes(x=Value)) +
  geom_histogram(bins=30, fill="#062e3e", alpha=0.7) +
  facet_wrap(~FinancialType, scales="free") +
  theme_minimal() +
  theme_economist() +
  scale_y_continuous(labels = scales::comma) +  # Using scales::comma format for y-axis
  labs(x='Value', y='Frequency', title='Distribution of Financial Details')
```


---


```{r warning=FALSE, include=FALSE}
# Decision tree
pacman::p_load(rpart.plot)

# Data manipulation
pacman::p_load(rgl, rattle, mice, dplyr, tidyverse)

# Plotting
pacman::p_load(viridis, hrbrthemes, ggplot2, heplots, ggpubr, forcats)
```


```{r warning=FALSE}
pacman::p_load(ranger)

# Identify the target variable and feature variables
target_variable <- "PotentialFraud"
feature_variables <- setdiff(names(data), target_variable)

# Define the formula for the random forest model
formula <- as.formula(paste(target_variable, "~ ."))

# Train the random forest model using ranger
rf_model <- ranger(formula,
                   data = data,
                   importance = 'impurity',  # This specifies to use Gini impurity for importance
                   num.trees = 100)

# Get the importance scores
importance_scores <- rf_model$variable.importance

# Create a data frame of the feature importance scores
importance_df <- data.frame(Feature = names(importance_scores), Importance = importance_scores)

# Order the features by importance
importance_df <- importance_df[order(-importance_df$Importance), ]
```

```{r warning=FALSE}
# Print the top 5 features
print("Top 10 features based on Gini importance:")
print(head(importance_df, 10))
```


```{r warning=FALSE, echo=FALSE}

# Visualize the top 10 features based on Gini importance
top_n_features <- head(importance_df, 10)

# Create a bar plot
ggplot(top_n_features, aes(x = reorder(Feature, Importance), y = Importance)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    coord_flip() +  # Flip coordinates for horizontal bars
    theme_minimal() +
    labs(title = "Top 10 Features Based on Gini Importance",
         x = "Features",
         y = "Importance Score") +
    theme(text = element_text(size = 12))
```




```{r warning=FALSE}
print(importance_df)
# shape of the data
dim(importance_df)
```


```{r warning=FALSE}
# Define a threshold for low importance; this could be a quantile, a fixed value, or based on domain knowledge
low_importance_threshold <- quantile(importance_df$Importance, 0.25)  # For example, the 25th percentile

# Identify the features that fall below this importance threshold
low_importance_features <- importance_df$Feature[importance_df$Importance <= low_importance_threshold]
print("Features with low importance:")
print(low_importance_features)
```



                                  Feature Importance
County                             County  13270.296
State                               State  11841.497
Log_OPTotalAmount       Log_OPTotalAmount  10319.958
Log_TotalClaimAmount Log_TotalClaimAmount   9116.225
Age                                   Age   8750.642





---


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(corrplot)

# List of numerical variables
numerical_vars <- c("Age", "Log_TotalClaimAmount", "Log_IPTotalAmount", "Log_OPTotalAmount", 
                    "UniquePhysCount")

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# print all the columns
names(data)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# Separate the data into numerical and categorical dataframes
numerical_data <- data[, numerical_vars]
categorical_data <- data[, categorical_vars]

# Perform Pearson correlation on numerical data
numerical_correlation <- cor(numerical_data, method = "pearson")
print("Numerical Correlation Matrix:")
print(numerical_correlation)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Visualize the numerical correlation matrix
corrplot(numerical_correlation, method = "color", addCoef.col = "black")
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
# save the data as csv
write.csv(data, "data.csv", row.names = FALSE)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# examine the data
str(data)
```